Exploring dimensions of biodiversity in Japanese ferns

### 2021-05-26 ms/manuscript.Rmd
Update MS

We checked for correlation between independent variables while taking into account spatial autocorrelation using a modified *t*-test with the `modified.ttest` in the `SpatialPack` R package. This indicated that temperature seasonality and % apomictic taxa were correlated with mean annual temperature. Therefore, we excluded temperature seasonality from modeling. We suspected that despite the correlation, % apomictic taxa might explain the distribution of genetic diversity better than temperature, so we conducted analyses separately for this variable as described below.

We constructed a linear model for each response variable (richness, SES of PD, SES of FD, SES of RPD, SES of RFD, PE p-score) dependent on all environmental variables (temperature, precipitation, and precipitation seasonality). We did not include any interaction terms as we had no a-priori reason to do so. Initial inspection of the relationship between richness and temperature showed a hump-shaped pattern, so we also included a quadratic term for temperature for the richness model only. We fit models including a Matérn correlation matrix based on the geographic centroids of each grid cell with the R package spaMM [@Rousset2014]. We verified the significance of each term of the model by conducting likelihood ratio tests (LRTs) between the full model and models each with a single term of interest removed, in sequence. We also verified the significance of the full model relative to the null model (only including the Matérn correlation matrix) using LRTs.

To compare between the models fit using temperature (environmental model) vs. % apomictic taxa (reproductive model) in response variables with a genetic component (SES of PD, SES of RPD, PE p-score), we constructed additional models where % apomictic taxa was used in place of temperature. We then conducted LRTs as described above, as well as between the environmental model vs. the reproductive model.

>>>>

We checked for correlation between independent variables while taking into account spatial autocorrelation using a modified *t*-test with the `modified.ttest` in the `SpatialPack` R package. This indicated that temperature seasonality and % apomictic taxa were correlated with mean annual temperature. The dataset including % apomictic taxa is smaller than the full dataset (_ vs _ taxa, respectively), so for the main modeling analysis we only included the following environmental variables: temperature, precipitation, and precipitation seasonality.

We constructed a linear model for each response variable (richness, SES of PD, SES of FD, SES of RPD, SES of RFD, PE p-score) dependent on the environmental variables. We did not include any interaction terms as we had no a-priori reason to do so. Initial inspection of the relationship between richness and temperature showed a hump-shaped pattern, so we also included a quadratic term for temperature for the richness model only. We fit models including a Matérn correlation matrix based on the geographic centroids of each grid cell as a random effect with the R package spaMM [@Rousset2014]. We verified the significance of each term of the model by conducting likelihood ratio tests (LRTs) between the full model and models each with a single term of interest removed, in sequence. We also verified the significance of the full model relative to the null model (only including the Matérn correlation matrix) using LRTs.

We postulated that % apomictic taxa might explain the distribution of genetic diversity better than temperature, so we also constructed a limited set of models to test this hypothesis. These models included response variables with a genetic component (SES of PD, SES of RPD, PE p-score) as dependent on either % apomictic taxa or temperature, precipitation, and precipitation seasonality, with random spatial effects as described above. We compared the fit between models including % apomictic taxa or temperature using conditional Akaike Information Criterion (cAIC).
